\huge UNet-HCRF Integration for EEG Pattern Recognition

### 2024-12-12 manuscript/main.tex
update background; fig position

\section{Background}
%Background & Related Work: background info and literature survey.

\sx{Add UNet}

Deep neural networks, particularly recurrent neural networks (RNNs) like LSTM,
are widely used for sequence labeling due to their ability to capture
long-range dependencies in sequences. Bidirectional LSTM (Bi-LSTM) further
enhances this by processing sequences in both directions, improving context
understanding.
CRFs, on the other hand, are commonly used for structured predictions,
modeling dependencies between neighboring labels to ensure valid output
sequences.
\citet{huang2015bidirectional} pioneered the use of Bi-LSTM integrated with
CRFs for sequence labeling tasks, demonstrating their effectiveness in
extracting features and ensuring label consistency.

However, Bi-LSTMs can still struggle to fully account for structural
dependencies between predicted labels, leading to potential inconsistencies in
the output sequence. Additionally, they typically require large amounts of
labeled data and are computationally expensive, making them challenging to
deploy in resource-constrained environments, such as hospitals.

Transformer-based models, such as BERT~\citep{devlin2019bert},
have emerged as powerful alternatives to traditional RNN-based approaches.
Trained on massive datasets in an unsupervised manner, Transformer can be
fine-tuned for various natural language processing tasks,
including sequence labeling.
\citet{devlin2019bert} directly compared BERT to traditional models, including
Bi-LSTM, and demonstrated its superior performance on tasks like named entity
recognition, highlighting its ability to capture more feature information
than LSTM models.

@@

@@

\begin{figure*}[tbp]
\centering
\includegraphics[width=\textwidth]{workframe}
\caption{\sx{caption}.}
\label{fig:workframe}
\end{figure*}

@@

@@

\begin{figure}[tbp]
\centering
\includegraphics[width=.48\textwidth]{example}
\caption{Examples of EEG patterns with different levels of expert agreement,
sourced from the HMS dataset.}
\label{fig:ex}
\end{figure}

@@

@@

@@

\begin{figure}[tbp]
\centering
\includegraphics[width=.45\textwidth]{inion}
\caption{Chains employed to determine representative montage features via
a bipolar montage, sourced from Kaggle.}
\label{fig:inion}
\end{figure}

\begin{figure}
\centering
\begin{subfigure}[b]{0.45\textwidth}
\includegraphics[width=\textwidth]{EEG_Signal}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.45\textwidth}
\includegraphics[width=\textwidth]{EEG_Spectrogram}
\end{subfigure}
\caption{An example of the denoised EEG signal (top) and its corresponding
spectrogram representation (bottom).}
\label{fig:EEG}
\end{figure}

@@

@@

\begin{figure}
\centering
\begin{subfigure}[b]{0.45\textwidth}
\includegraphics[width=\textwidth]{pretrain}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.45\textwidth}
\includegraphics[width=\textwidth]{finetune}
\end{subfigure}
\caption{The learning rate schedulers for the pre-trained UNet (top) and
the fine-tuned HCRF (bottom).}
\label{fig:cos}
\end{figure}

>>>>

\section{Background}

Deep neural networks, particularly convolutional architectures like U-net, have 
proven highly effective for image segmentation tasks due to their ability to
capture rich spatial features and contextual information at different scales.
U-net, in particular, has demonstrated remarkable performance in medical image
segmentation by leveraging its encoder-decoder architecture with skip
connections to preserve spatial resolution and improve
localization~\citep{ronneberger2015unet,zhou2018unet}.

Traditional sequence-labeling approaches, such as recurrent neural networks
(RNNs) like LSTM and bidirectional LSTMs (Bi-LSTM), have been effective in
capturing long-range dependencies and processing temporal information. However,
these models are less suited for spatially rich tasks like image segmentation.
Moreover, LSTMs and Bi-LSTMs often require large labeled datasets, are
computationally expensive, and struggle to account for the structural
dependencies present in spatial data, which limits their applicability in
resource-constrained environments like hospitals.

Transformer-based models, such as BERT, have emerged as powerful tools for
sequence-labeling tasks, demonstrating superior performance over traditional
RNN-based models. However, their advantages are primarily realized in natural
language processing rather than spatially dependent tasks like image
segmentation.

U-net addresses these challenges by focusing on capturing structural
dependencies in spatial domains. For EEG data, where the objective shifts from
sequence-labeling to segmentation, U-net is a more appropriate choice. Its
ability to efficiently learn and propagate contextual and spatial information
makes it a robust alternative to sequence-based methods like Bi-LSTM.

@@

\begin{figure*}[ht]
\centering
\includegraphics[width=\textwidth]{workframe}
\caption{The proposed feature-learning-based HCRF using UNet feature maps.}
\label{fig:workframe}
\end{figure*}

Figure~\ref{fig:workframe} illustrates the descriptive workframe layout
employed in this study, highlighting the key components and processes involved.

@@

@@

\begin{figure}[ht]
\centering
\includegraphics[width=.48\textwidth]{example}
\caption{Examples of EEG patterns with different levels of expert agreement,
sourced from the HMS dataset.}
\label{fig:ex}
\end{figure}

@@

@@

\begin{figure}[bp]
\centering
\includegraphics[width=.45\textwidth]{inion}
\caption{Chains employed to determine representative montage features via
a bipolar montage, sourced from Kaggle.}
\label{fig:inion}
\end{figure}

@@

\begin{figure}[tp]
\centering
\begin{subfigure}[b]{0.45\textwidth}
\includegraphics[width=\textwidth]{EEG_Signal}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.45\textwidth}
\includegraphics[width=\textwidth]{EEG_Spectrogram}
\end{subfigure}
\caption{An example of the denoised EEG signal (top) and its corresponding
spectrogram representation (bottom).}
\label{fig:EEG}
\end{figure}

@@

@@

\begin{figure}[tbp]
\centering
\begin{subfigure}[b]{0.45\textwidth}
\includegraphics[width=\textwidth]{pretrain}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.45\textwidth}
\includegraphics[width=\textwidth]{finetune}
\end{subfigure}
\caption{The learning rate schedulers for the pre-trained UNet (top) and
the fine-tuned HCRF (bottom).}
\label{fig:cos}
\end{figure}

@@
